---
title: "LocalFact Data Analysis"
author: "Sofia Bliss-Carrascoas"
date: "3/10/2023"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages, warning=FALSE, message=FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
library(tibble)
library(kableExtra)
library(lubridate)
library(ggplot2)
```

```{r loaddata, warning=FALSE, message=FALSE, echo = FALSE }
LocalFact <- read_csv("LocalFactDataCleaned.csv")

LocalFact <- LocalFact %>%
  mutate(Date = mdy(Date))

LocalFact <- LocalFact %>%
  mutate(Month = months(Date))

LocalFact$Month <- factor(LocalFact$Month, 
                              levels = c("January", 
                                         "February", 
                                         "March",
                                         "April",
                                         "May",
                                         "June",
                                         "July",
                                         "August",
                                         "September",
                                         "October",
                                         "November"))
```
Basic Queries:
number of total checked claims
```{r}
LocalFact %>%
  count()
```

number of unique fact-checks published
```{r}
LocalFact %>%
  distinct(`Fact-Check URL`) %>%
  count()
```

number of fact-checks by state
```{r state, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  group_by(State) %>%
  count()
```

number of fact-checks by fact-checker (organization)
```{r org, warning=FALSE, message=FALSE, echo = FALSE}
OrgCount <- LocalFact %>%
  group_by(Organization) %>%
  count()

OrgCount %>%
  arrange(desc(n))

OrgCount %>%
  summary(n)

OrgCount %>%
  ggplot(aes(x = n)) + 
  geom_boxplot() +
  labs(title = "Frequency of Output by Organization",
       x = "Total Number of Fact Checks")
```

number of fact-checks by political/non-political type of claimant
```{r type, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  group_by(`Type of Claimant`) %>%
  count() 
```

number of fact-checks by month of the year
```{r month, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  group_by(Month) %>%
  count() %>%
  ggplot(aes(x = Month, y = n)) +
  geom_point() +
  labs(title = "Number of Fact-Checked Claims by Month", y = "Count")
```

number of fact-checks by politician (claimant name + politician as category)
```{r pol, warning=FALSE, message=FALSE, echo = FALSE}
Politician <- LocalFact %>%
  filter(`Political Designation` == "Politician") %>%
  group_by(`Claimant Name // (Claim Source?)`) %>%
  count()

Politician %>%
  arrange(desc(n))

Politician %>%
  summary(n)

Politician %>%
  ggplot(aes(x = n)) + 
  geom_boxplot() +
  labs(title = "How much were politicians singled out?",
       x = "Total Number of Fact Checks per Politician")
```

number of fact-checks by re-occurring non-political types of claimant
```{r nonpoltype, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  filter(`Claimant Name // (Claim Source?)` == "Viewer question" | 
           `Claimant Name // (Claim Source?)` == "You may have heard..." |
           `Claimant Name // (Claim Source?)` == "Unknown social media user(s)") %>%
  group_by(`Claimant Name // (Claim Source?)`) %>%
  count() 
```

number of fact-checks by incumbent seeking re-election (same office held as office sought)
```{r incumbent, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  filter(`Aspiring Position` == `Current Position`) %>%
  group_by(`Claimant Name // (Claim Source?)`) %>%
  count() %>%
  arrange(desc(n))
```

number of fact-checks by current officeholder (yes)
```{r officeholder, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  filter(`Current Officeholder?` == "Yes") %>%
  group_by(`Claimant Name // (Claim Source?)`) %>%
  count() %>%
  arrange(desc(n))
```

number of fact-checks by running for office (yes)
```{r candidate, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  filter(`Running for Office?` == "Yes") %>%
  group_by(`Claimant Name // (Claim Source?)`) %>%
  count() %>%
  arrange(desc(n))
```

number of fact-checks by office held
```{r office, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  filter(!is.na(`Current Position`)) %>%
  group_by(`Current Position`) %>%
  count() %>%
  arrange(desc(n))
```

number of fact-checks by office sought
```{r officesought, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  filter(!is.na(`Aspiring Position`)) %>%
  group_by(`Aspiring Position`) %>%
  count() %>%
  arrange(desc(n))
```

number of fact-checks by topic
```{r topictag, warning=FALSE, message=FALSE, echo = FALSE}
TOPICTAGS <- LocalFact %>%
  summarise(`INDIVIDUAL RECORD` = sum(`INDIVIDUAL RECORD`),
            `ECONOMIC ISSUES` = sum(`ECONOMIC ISSUES`),
            `GOVERNMENT OPERATIONS` = sum(`GOVERNMENT OPERATIONS`),
            `VOTING/ELECTIONS` = sum(`VOTING/ELECTIONS`),
            `CIVIL RIGHTS MINORITY ISSUES AND CIVIL LIBERTIES` = sum(`CIVIL RIGHTS MINORITY ISSUES AND CIVIL LIBERTIES`),
            `IMMIGRATION AND REFUGEE ISSUES` = sum(`IMMIGRATION AND REFUGEE ISSUES`),
            `LGBTQ` = sum(`LGBTQ`),
            `RACE` = sum(`RACE`),
            `RELIGION` = sum(`RELIGION`),
            `SOCIAL WELFARE` = sum(`SOCIAL WELFARE`),
            `EDUCATION` = sum(`EDUCATION`),
            `ABORTION/WOMEN` = sum(`ABORTION/WOMEN`),
            `HEALTH` = sum(`HEALTH`),
            `LAW AND CRIME AND POLICING` = sum(`LAW AND CRIME AND POLICING`),
            `GUNS` = sum(`GUNS`),
            `HISTORY` = sum(`HISTORY`),
            `LEGISLATION` = sum(`LEGISLATION`),
            `AGRICULTURE` = sum(`AGRICULTURE`),
            `ENERGY AND ENVIRONMENT` = sum(`ENERGY AND ENVIRONMENT`),
            `COMMUNITY DEVELOPMENT AND HOUSING` = sum(`COMMUNITY DEVELOPMENT AND HOUSING`),
            `TECHNOLOGY SCIENCE SPACE` = sum(`TECHNOLOGY SCIENCE SPACE`),
            `MEDIA AND COMMUNICATIONS` = sum(`MEDIA AND COMMUNICATIONS`))
DATATOPICTAGS <- t(TOPICTAGS)
DATATOPICTAGS <- as.data.frame(DATATOPICTAGS)
DATATOPICTAGS$Tag <- rownames(DATATOPICTAGS)
DATATOPICTAGS %>%
  arrange(desc(V1))

write.csv(DATATOPICTAGS, "LocalFactTopicTags.csv", row.names=FALSE)
```

number of fact-checks by state and topic
```{r statetopic, warning=FALSE, message=FALSE, echo = FALSE}
STATETOPICTAGS <- LocalFact %>%
  group_by(State) %>%
  summarise(`INDIVIDUAL RECORD` = sum(`INDIVIDUAL RECORD`),
            `ECONOMIC ISSUES` = sum(`ECONOMIC ISSUES`),
            `GOVERNMENT OPERATIONS` = sum(`GOVERNMENT OPERATIONS`),
            `VOTING/ELECTIONS` = sum(`VOTING/ELECTIONS`),
            `CIVIL RIGHTS MINORITY ISSUES AND CIVIL LIBERTIES` = sum(`CIVIL RIGHTS MINORITY ISSUES AND CIVIL LIBERTIES`),
            `IMMIGRATION AND REFUGEE ISSUES` = sum(`IMMIGRATION AND REFUGEE ISSUES`),
            `LGBTQ` = sum(`LGBTQ`),
            `RACE` = sum(`RACE`),
            `RELIGION` = sum(`RELIGION`),
            `SOCIAL WELFARE` = sum(`SOCIAL WELFARE`),
            `EDUCATION` = sum(`EDUCATION`),
            `ABORTION/WOMEN` = sum(`ABORTION/WOMEN`),
            `HEALTH` = sum(`HEALTH`),
            `LAW AND CRIME AND POLICING` = sum(`LAW AND CRIME AND POLICING`),
            `GUNS` = sum(`GUNS`),
            `HISTORY` = sum(`HISTORY`),
            `LEGISLATION` = sum(`LEGISLATION`),
            `AGRICULTURE` = sum(`AGRICULTURE`),
            `ENERGY AND ENVIRONMENT` = sum(`ENERGY AND ENVIRONMENT`),
            `COMMUNITY DEVELOPMENT AND HOUSING` = sum(`COMMUNITY DEVELOPMENT AND HOUSING`),
            `TECHNOLOGY SCIENCE SPACE` = sum(`TECHNOLOGY SCIENCE SPACE`),
            `MEDIA AND COMMUNICATIONS` = sum(`MEDIA AND COMMUNICATIONS`))

STATETOPICTAGS
write.csv(STATETOPICTAGS, "LocalFactSTATETopicTags.csv", row.names=FALSE)

```

number of fact-checks by state and topic
```{r poltopic, warning=FALSE, message=FALSE, echo = FALSE}
POLITICIANTOPICTAGS <- LocalFact %>%
  filter(`Political Designation` == "Politician") %>%
  summarise(`INDIVIDUAL RECORD` = sum(`INDIVIDUAL RECORD`),
            `ECONOMIC ISSUES` = sum(`ECONOMIC ISSUES`),
            `GOVERNMENT OPERATIONS` = sum(`GOVERNMENT OPERATIONS`),
            `VOTING/ELECTIONS` = sum(`VOTING/ELECTIONS`),
            `CIVIL RIGHTS MINORITY ISSUES AND CIVIL LIBERTIES` = sum(`CIVIL RIGHTS MINORITY ISSUES AND CIVIL LIBERTIES`),
            `IMMIGRATION AND REFUGEE ISSUES` = sum(`IMMIGRATION AND REFUGEE ISSUES`),
            `LGBTQ` = sum(`LGBTQ`),
            `RACE` = sum(`RACE`),
            `RELIGION` = sum(`RELIGION`),
            `SOCIAL WELFARE` = sum(`SOCIAL WELFARE`),
            `EDUCATION` = sum(`EDUCATION`),
            `ABORTION/WOMEN` = sum(`ABORTION/WOMEN`),
            `HEALTH` = sum(`HEALTH`),
            `LAW AND CRIME AND POLICING` = sum(`LAW AND CRIME AND POLICING`),
            `GUNS` = sum(`GUNS`),
            `HISTORY` = sum(`HISTORY`),
            `LEGISLATION` = sum(`LEGISLATION`),
            `AGRICULTURE` = sum(`AGRICULTURE`),
            `ENERGY AND ENVIRONMENT` = sum(`ENERGY AND ENVIRONMENT`),
            `COMMUNITY DEVELOPMENT AND HOUSING` = sum(`COMMUNITY DEVELOPMENT AND HOUSING`),
            `TECHNOLOGY SCIENCE SPACE` = sum(`TECHNOLOGY SCIENCE SPACE`),
            `MEDIA AND COMMUNICATIONS` = sum(`MEDIA AND COMMUNICATIONS`))

POLITICIANTOPICTAGS$`Subset of Data` <- "Politician"

POLITICALORGTOPICTAGS <- LocalFact %>%
  filter(`Political Designation` == "Political organization") %>%
  summarise(`INDIVIDUAL RECORD` = sum(`INDIVIDUAL RECORD`),
            `ECONOMIC ISSUES` = sum(`ECONOMIC ISSUES`),
            `GOVERNMENT OPERATIONS` = sum(`GOVERNMENT OPERATIONS`),
            `VOTING/ELECTIONS` = sum(`VOTING/ELECTIONS`),
            `CIVIL RIGHTS MINORITY ISSUES AND CIVIL LIBERTIES` = sum(`CIVIL RIGHTS MINORITY ISSUES AND CIVIL LIBERTIES`),
            `IMMIGRATION AND REFUGEE ISSUES` = sum(`IMMIGRATION AND REFUGEE ISSUES`),
            `LGBTQ` = sum(`LGBTQ`),
            `RACE` = sum(`RACE`),
            `RELIGION` = sum(`RELIGION`),
            `SOCIAL WELFARE` = sum(`SOCIAL WELFARE`),
            `EDUCATION` = sum(`EDUCATION`),
            `ABORTION/WOMEN` = sum(`ABORTION/WOMEN`),
            `HEALTH` = sum(`HEALTH`),
            `LAW AND CRIME AND POLICING` = sum(`LAW AND CRIME AND POLICING`),
            `GUNS` = sum(`GUNS`),
            `HISTORY` = sum(`HISTORY`),
            `LEGISLATION` = sum(`LEGISLATION`),
            `AGRICULTURE` = sum(`AGRICULTURE`),
            `ENERGY AND ENVIRONMENT` = sum(`ENERGY AND ENVIRONMENT`),
            `COMMUNITY DEVELOPMENT AND HOUSING` = sum(`COMMUNITY DEVELOPMENT AND HOUSING`),
            `TECHNOLOGY SCIENCE SPACE` = sum(`TECHNOLOGY SCIENCE SPACE`),
            `MEDIA AND COMMUNICATIONS` = sum(`MEDIA AND COMMUNICATIONS`))

POLITICALORGTOPICTAGS$`Subset of Data` <- "Political organization"

GENTOPICTAGS <- TOPICTAGS

GENTOPICTAGS$`Subset of Data` <- "All Data"

COMPARETOPICTAGS <- full_join(POLITICIANTOPICTAGS, POLITICALORGTOPICTAGS, match = all)

COMPARETOPICTAGS <- full_join(COMPARETOPICTAGS, GENTOPICTAGS, match = all)

COMPARETOPICTAGS
```

number of fact-checks by definitive rating
```{r defrating, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  group_by(`Was a Definitive Rating Given?`) %>%
  count()
```

number of fact-checks by source of statement
```{r source, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  group_by(`Source of Statement`) %>%
  count() %>%
  arrange(desc(n))
```

number of fact-checks by fact-check format
```{r format, warning=FALSE, message=FALSE, echo = FALSE}
LocalFact %>%
  group_by(`Fact-Check Format`) %>%
  count() %>%
  arrange(desc(n))
```

what percentage had multiple claims?
```{r splits, warning=FALSE, message=FALSE, echo = FALSE}
uniqueclaims <- LocalFact %>%
  distinct(`Fact-Check URL`) %>%
  count()

splitclaims <- LocalFact %>%
  filter(`Multiple claims?` == "A") %>%
  count()

splitclaims/uniqueclaims*100


```


