---
title: "LocalFact Data Analysis"
author: "Sofia Bliss-Carrascoas"
date: "3/10/2023"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages, warning=FALSE, message=FALSE, echo = FALSE}
library(tidyverse)
library(knitr)
library(tibble)
library(kableExtra)
library(lubridate)
```

```{r color scale and ordering, warning=FALSE, message=FALSE, echo = FALSE }
LocalFact <- read_csv("midedit data.csv")

LocalFact %>%
  filter(is.na(State))
```
Here are some of the things I've jotted down, if it's helpful:
number of fact-checks by state
```{r}
LocalFact %>%
  group_by(State) %>%
  count() %>%
  kable()
```

number of fact-checks by fact-checker (organization)
```{r}
LocalFact %>%
  group_by(Organization) %>%
  count() %>%
  kable()
```

number of fact-checks by political/non-political type of claimant
```{r}

```

number of fact-checks by month of the year
```{r}
LocalFact <- LocalFact %>%
  mutate(Date = mdy(Date))

LocalFact <- LocalFact %>%
  mutate(Month = months(Date))

LocalFact %>%
  group_by(Month) %>%
  count() %>%
  kable()
```

number of fact-checks by politician (claimant name + politician as category)
```{r}

```

number of fact-checks by incumbent seeking re-election (same office held as office sought)
```{r}

```

number of fact-checks by current officeholder (yes)
```{r}

```

number of fact-checks by running for office (yes)
```{r}

```

number of fact-checks by office held
```{r}

```

number of fact-checks by office sought
```{r}

```

number of fact-checks by topic
```{r}

```

number of fact-checks by state and topic
```{r}

```

number of fact-checks by definitive rating
```{r}

```

number of fact-checks by source of statement
```{r}

```

number of fact-checks by fact-check format
```{r}

```

what percentage had multiple claims?
```{r}

```


